---
title: "PEP Scheduling"
format: html
editor: visual
---

This report was generated on `r Sys.Date()`.

## Introduction

The figures in this document represent the upcoming work for PEP staff over the next 60 days. The intent of this information is to help support cross-project planning and communication

## Planned Work by Action Lead

NOTE: For actions where there is no start date assigned, an anticipated start date of 60 days before the planned end date has been estimated.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(RPostgreSQL)
library(ggplot2)
library(scales)

# Connect to DB and get starting data
con <- RPostgreSQL::dbConnect(PostgreSQL(), 
                              dbname = Sys.getenv("pep_db"), 
                              host = Sys.getenv("pep_ip"), 
                              user = Sys.getenv("pep_admin"), 
                              password =  Sys.getenv("admin_pw"))

planning <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling
                                    WHERE (planned_end_date <= (now() + \'60 days\'::interval)
                                    OR (planned_start_date <= now() AND planned_end_date >= now()))
                                    AND activity_lead IS NOT NULL") %>%
  mutate(adjusted_start = as.Date(ifelse(planned_start_date < Sys.Date() - 30, Sys.Date() - 30, planned_start_date), origin = '1970-01-01'),
         adjusted_end = as.Date(ifelse(planned_end_date > Sys.Date() + 60, Sys.Date() + 60, 
                                       ifelse(planned_end_date < Sys.Date(), Sys.Date() + 60, planned_end_date)), origin = '1970-01-01'),
         ylab_staff = str_wrap(paste(activity_plan_nickname, activity_name, sep = ": "), width = 40),
         ylab_project = str_wrap(paste0(activity_plan_nickname,": ", activity_name, " (", activity_lead, ")"), width = 40))

staff <- unique(planning$activity_lead)
activityPlan <- planning %>%
  arrange(activity_plan_nickname) %>% 
  select(activity_plan_nickname) %>%
  unique()
  
```

```{r figures_byStaff, results = 'all', fig.height = 7, fig.width = 10}
# Figure 1
for (i in 1:length(staff)) {
  figure_staff <- planning %>%
    filter(activity_lead == staff[i]) %>%
    mutate(ylab = factor(ylab_staff, levels = rev(sort(ylab_staff))))
  
  fig <- ggplot(figure_staff, aes(x = adjusted_start, xend = adjusted_end, 
                     y = ylab, yend = ylab, 
                     color = activity_status)) +
    geom_segment(size = 5) +
    scale_color_manual(values = c("Not yet started" = "darkslategray",
                                "Awaiting dependency" = "darkgoldenrod2",
                                "In progress" = "chartreuse4")) + 
    scale_x_date(limits = c(Sys.Date() - 31, Sys.Date() + 180),
                 date_labels = "%Y-%m-%d", 
                 date_breaks ="1 month") + 
    labs(title = paste("Scheduled Actions for ", staff[i], sep = ""),
         x = "Date",
         y = "Action",
         color = "Action Status") +
    theme(panel.grid.minor = element_blank()) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
  
  print(fig)  
}
```

```{r figures_byProject, results = 'all', fig.height = 7, fig.width = 10}
# Figure 2
for (a in 1:nrow(activityPlan)) {
  print(activityPlan$activity_plan_nickname[a]) 
  
  component <- planning %>%
    filter(activity_plan_nickname == activityPlan$activity_plan_nickname[a]) %>%
    select(component_name) %>%
    arrange(component_name) %>%
    unique() 
  
  for (c in 1:nrow(component)) {
    print(component$component_name[c])
    
    project <- planning %>% 
      filter(activity_plan_nickname == activityPlan$activity_plan_nickname[a]) %>%
      filter(component_name == component$component_name[c]) %>% 
      select(project_name) %>%
      arrange(project_name) %>%
      unique()
    
    for (p in 1:nrow(project)) {
      figure_project <- planning %>% 
        filter(activity_plan_nickname == activityPlan$activity_plan_nickname[a]) %>%
        filter(component_name == component$component_name[c]) %>% 
        filter(project_name == project$project_name[p]) %>% 
        mutate(ylab = factor(ylab_project, levels = rev(sort(ylab_project))))
      
      fig <- ggplot(figure_project, aes(x = adjusted_start, xend = adjusted_end,
                                       y = ylab, yend = ylab, 
                                       color = activity_status,
                                       label = activity_lead)) +
        geom_segment(size = 5) +
        #geom_label(hjust = 0) +
        scale_color_manual(values = c("Not yet started" = "darkslategray",
                            "Awaiting dependency" = "darkgoldenrod2",
                            "In progress" = "chartreuse4")) +
        scale_x_date(limits = c(Sys.Date() - 31, Sys.Date() + 180),
             date_labels = "%Y-%m-%d", 
             date_breaks ="1 month") + 
        labs(title = project$project_name[p],
             x = "Date",
             y = "Action",
             color = "Action Status") +
        theme(panel.grid.minor = element_blank()) +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
      
      print(fig)
    }
  }
}

```
