---
title: "PEP Scheduling Report"
toc: true
theme: superhero
format: html
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(RPostgreSQL)
library(ggplot2)
library(scales)
library(ggpattern)
library(stringr)

# Connect to DB and get starting data
con <- RPostgreSQL::dbConnect(PostgreSQL(), 
                              dbname = Sys.getenv("mml_pep_db"), 
                              host = Sys.getenv("mml_pep_ip"), 
                              user = Sys.getenv("mml_pep_user"), 
                              password =  Sys.getenv("mml_pep_user_pw"))

priorities <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.qry_project_priorities") 

p_critical <- priorities %>%
  filter(project_priority == "Mission Critical") %>%
  select(-project_priority) %>%
  arrange(activity_plan_nickname, project_name) %>%
  rename("Activity Plan" = activity_plan_nickname,
         "Project" = project_name,
         "Status" = project_status)

p_high <- priorities %>%
  filter(project_priority == "High Priority") %>%
  select(-project_priority) %>%
  arrange(activity_plan_nickname, project_name) %>%
  rename("Activity Plan" = activity_plan_nickname,
         "Project" = project_name,
         "Status" = project_status)

qaqc <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.qaqc_roles_missing_info") %>%
  rename("Staff" = last_name,
         "# Missing Info" = num_missing_info)

work <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling") 

# work00 <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling_00days") %>%
#   mutate(work_intensity = ifelse(is.na(work_intensity), "Unknown", work_intensity))  %>%
#   mutate(work_intensity = factor(work_intensity, levels=c("High", "Moderate", "Low", "Unknown"), ordered = TRUE)) %>%
#   mutate(num_roles = ifelse(num_roles == 1, "1", 
#                             ifelse(num_roles == 2 | num_roles == 3, "2-3",
#                                    ifelse(num_roles >=4 & num_roles <= 5, "4-5",
#                                           ifelse(num_roles >= 6 & num_roles <= 8, "6-8", "9+")))))
#   
# work30 <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling_30days") %>%
#   mutate(work_intensity = ifelse(is.na(work_intensity), "Unknown", work_intensity))  %>%
#   mutate(work_intensity = factor(work_intensity, levels=c("High", "Moderate", "Low", "Unknown"), ordered = TRUE)) %>%
#   mutate(num_roles = ifelse(num_roles == 1, "1", 
#                             ifelse(num_roles == 2 | num_roles == 3, "2-3",
#                                    ifelse(num_roles >=4 & num_roles <= 5, "4-5",
#                                           ifelse(num_roles >= 6 & num_roles <= 8, "6-8", "9+")))))
# 
# work90 <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling_90days") %>%
#   mutate(work_intensity = ifelse(is.na(work_intensity), "Unknown", work_intensity))  %>%
#   mutate(work_intensity = factor(work_intensity, levels=c("High", "Moderate", "Low", "Unknown"), ordered = TRUE)) %>%
#   mutate(num_roles = ifelse(num_roles == 1, "1", 
#                             ifelse(num_roles == 2 | num_roles == 3, "2-3",
#                                    ifelse(num_roles >=4 & num_roles <= 5, "4-5",
#                                           ifelse(num_roles >= 6 & num_roles <= 8, "6-8", "9+")))))

work_manuscripts <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling_manuscripts") 
manuscripts <- unique(work_manuscripts$project_name) 

work_reports <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling_reports") 
reports <- unique(work_reports$project_name) 

work_fy <- RPostgreSQL::dbGetQuery(con, "SELECT * FROM administrative.summ_4scheduling_fy") %>%
  mutate(pp = ifelse(performance_plan_element == 0, project_pp_element, performance_plan_element)) %>%
  filter(!is.na(pp)) %>%
  mutate(performance_plan_element = factor(performance_plan_element, levels = c(1, 2, 3, 4, 5), 
                                           labels=c("Element 1", "Element 2", "Element 3", "Element 4", "Element 5"), ordered = TRUE))

work_fy_priorities <- work_fy %>%
  filter(project_priority != 'Other')

RPostgreSQL::dbDisconnect(con)
due_soon <- work %>%
  filter(planned_end_date > Sys.Date() & planned_end_date <= Sys.Date() + 10) %>%
  select(project_activity, role_name, staff, planned_end_date) %>%
  rename("Project" = project_activity,
         "Role" = role_name,
         "Staff" = staff,
         "Planned End Date" = planned_end_date)

overdue <- work %>%
  filter(planned_end_date <= Sys.Date()) %>%
  select(project_activity, role_name, staff, planned_end_date) %>%
  rename("Project" = project_activity,
         "Role" = role_name,
         "Staff" = staff,
         "Planned End Date" = planned_end_date)

planning <- work %>%
  filter(planned_start_date <= Sys.Date() + 150 & planned_end_date > Sys.Date() - 30) %>%
  mutate(adjusted_start = as.Date(ifelse(planned_start_date < Sys.Date() - 30, Sys.Date() - 30, planned_start_date), origin = '1970-01-01'),
         adjusted_end = as.Date(ifelse(planned_end_date > Sys.Date() + 150, Sys.Date() + 150, planned_end_date), origin = '1970-01-01'),
         ylab_staff = str_wrap(paste(project_activity, role_name, sep = ": "), width = 75),
         ylab_project = str_wrap(paste0(project_activity,": ", role_name, " (", staff, ")"), width = 75),
         work_intensity = ifelse(is.na(work_intensity), "Unknown", work_intensity))  %>%
  mutate(performance_plan_element = factor(performance_plan_element, levels = c(0, 1, 2, 3, 4, 5), 
                                           labels=c("Project lead", "Elem#1", "Elem#2", "Elem#3", "Elem#4", "Elem5"), ordered = TRUE)) %>%
  mutate(work_intensity = factor(work_intensity, levels=c("High", "Moderate", "Low", "Unknown"), ordered = TRUE)) %>%
  mutate(project_priority = factor(project_priority, levels=c("Mission Critical", "High Priority", "Other", "Unknown"), ordered = TRUE)) %>%
  mutate(duration = adjusted_end - adjusted_start)
           

pep <- unique(planning$staff) 

activityPlan <- planning %>%
  arrange(activity_plan_nickname) %>% 
  select(activity_plan_nickname) %>%
  unique()
```

This report was generated on `r Sys.Date()`.

# Introduction

The information in this report represents the outstanding and upcoming work for PEP staff over the next 180 days. The intent of this information is to help support cross-project planning and communication.

# Project Priorities

The following tables list projects by their priority level.

<details>

<summary>Click to Expand</summary>

::: panel-tabset
## Mission Critical Projects
```{r mission_critical, results='asis'}
knitr::kable(p_critical, format = "pipe", padding = 2)
```

## High Priority Projects
```{r high_priority, results='asis'}
knitr::kable(p_high, format = "pipe", padding = 2)
```
:::

</details>

# Roles Missing Info

The following table lists the the number of roles for each staff that are missing required information in the DB.

<details>

<summary>Click to Expand</summary>

```{r qaqc, results='asis'}
if (nrow(qaqc) == 0) {
  cat("\n Hooray!! All records are complete!\n ")
} else {
  knitr::kable(qaqc, format = "pipe", padding = 2)
}
```

</details>

# Work Due Soon

The following table lists the roles that are expected to end within the next 10 days in the DB.

```{r due_soon, results='asis'}
if (nrow(due_soon) == 0) {
  cat("\n Hooray!! No upcoming deadlines!\n ")
} else {
  knitr::kable(due_soon, format = "pipe", padding = 2)
}
```

# Overdue Work

The following table lists the roles that are overdue in the DB.

```{r overdue_work, results='asis'}
if (nrow(overdue) == 0) {
  cat("\n Hooray!! No overdue work!\n ")
} else {
  knitr::kable(overdue, format = "pipe", padding = 2)
}
```

<!-- # Potential Bottlenecks -->

<!-- The following figures display the number of roles by work intensity by staff for this week, in the next 30 days and in the next 90 days. -->

<!-- ::: panel-tabset -->
<!-- ## This Week -->

<!-- ```{r bottleneck00} -->
<!-- ggplot(work00, aes(work_intensity, ordered(staff), fill = num_roles)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_manual("# Roles", -->
<!--                     values = c("1" = "darkolivegreen3", -->
<!--                                   "2-3" = "gold", -->
<!--                                   "4-5" = "darkorange3", -->
<!--                                   "6-8" = "firebrick4", -->
<!--                                   "9+" = "orchid4")) + -->
<!--   labs(title = "# of Roles by Intensity - This Week", -->
<!--      x = "Work Intensity", -->
<!--      y = "Staff") -->
<!-- ``` -->

<!-- ## Next 30 Days -->

<!-- ```{r bottleneck30} -->
<!-- ggplot(work30, aes(work_intensity, ordered(staff), fill = num_roles)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_manual("# Roles", -->
<!--                     values = c("1" = "darkolivegreen3", -->
<!--                                   "2-3" = "gold", -->
<!--                                   "4-5" = "darkorange3", -->
<!--                                   "6-8" = "firebrick4", -->
<!--                                   "9+" = "orchid4")) + -->
<!--   labs(title = "# of Roles by Intensity - 30 Days", -->
<!--      x = "Work Intensity", -->
<!--      y = "Staff") -->
<!-- ``` -->

<!-- ## Next 90 Days -->

<!-- ```{r bottleneck90} -->
<!-- ggplot(work90, aes(work_intensity, ordered(staff), fill = num_roles)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_manual("# Roles", -->
<!--                     values = c("1" = "darkolivegreen3", -->
<!--                                   "2-3" = "gold", -->
<!--                                   "4-5" = "darkorange3", -->
<!--                                   "6-8" = "firebrick4", -->
<!--                                   "9+" = "orchid4")) + -->
<!--   labs(title = "# of Roles by Intensity - 90 Days", -->
<!--      x = "Work Intensity", -->
<!--      y = "Staff") -->
<!-- ``` -->
<!-- ::: -->

# Planned Work by PEP Staff

Today's date is shown as a grey vertical line on the figure. Roles that are missing "work intensity" appear on the y-axis of each chart, but the start/end dates are not visible.

```{r figures_byStaff, results = 'asis', fig.width = 10, fig.height = 8}
cat("::: {.panel-tabset}")
cat("\n")

for (i in 1:length(pep)) {
  cat("\n\n## ", pep[i], "\n")

  figure_staff <- planning %>%
    filter(staff == pep[i]) %>% 
    filter(!(duration >= 180 & work_intensity == 'Low')) %>%
    mutate(ylab = factor(ylab_staff, levels = rev(sort(ylab_staff)))) %>%
    arrange(project_activity, adjusted_end)

  fig <- ggplot(figure_staff, aes(x = adjusted_start, xend = adjusted_end, y = ylab, yend = ylab,
                                  color = performance_plan_element, alpha = project_priority)) +
    geom_vline(aes(xintercept = Sys.Date()), col = "darkgray", linewidth = 1.5) +
    geom_segment(size = 5) +
    scale_color_manual(name = "Performance Plan",
                       values = c("Project lead" = "deeppink4",
                                  "Elem#1" = "chocolate1",
                                  "Elem#2" = "darkgoldenrod2",
                                  "Elem#3" = "darkolivegreen4",
                                  "Elem#4" = "cadetblue",
                                  "Elem#5" = "orchid4")) +
      scale_alpha_manual(name = "Project Priority",
                       values = c("Mission Critical" = 1,
                                  "High Priority" = 0.45,
                                  "Other" = 0.1,
                                  "Unknown" = 0)) +
    # scale_alpha_manual(name = "Work Intensity", # Using work intensity field
    #                    values = c("High" = 1,
    #                               "Moderate" = 0.45,
    #                               "Low" = 0.1,
    #                               "Unknown" = 0)) +
    scale_x_date(limits = c(Sys.Date() - 30, Sys.Date() + 150),
                 date_labels = "%Y-%m-%d",
                 date_breaks ="1 month") +
    labs(title = paste("Scheduled Roles for ", pep[i], sep = ""),
         x = "Date",
         y = "Role") +
    theme(plot.title.position = "plot") +
    theme(panel.grid.minor = element_blank()) +
    theme(axis.text.x = element_text(vjust = 0.5, hjust = 1)) +
    theme(legend.position = "top", legend.box = "vertical", legend.margin = margin()) 

  print(fig)
  cat("\n")
}
cat(":::")
```

# Planned Work by Project

Today's date is shown as a grey vertical line on the figure. Roles that are missing "work intensity" appear on the y-axis of each chart, but the start/end dates are not visible.

<details>

<summary>Click to Expand</summary>

```{r figures_byProject, results = 'asis', fig.width = 10}
for (a in 1:nrow(activityPlan)) {
  cat("\n\n## ", activityPlan$activity_plan_nickname[a], "\n")

  component <- planning %>%
    filter(activity_plan_nickname == activityPlan$activity_plan_nickname[a]) %>%
    select(component_name) %>%
    arrange(component_name) %>%
    unique()

  for (c in 1:nrow(component)) {
   cat("\n")
   cat("\n\n### ", component$component_name[c], "\n")

    project <- planning %>%
      filter(activity_plan_nickname == activityPlan$activity_plan_nickname[a]) %>%
      filter(component_name == component$component_name[c]) %>%
      select(project_name) %>%
      arrange(project_name) %>%
      unique()

    for (p in 1:nrow(project)) {
      figure_project <- planning %>%
        filter(activity_plan_nickname == activityPlan$activity_plan_nickname[a]) %>%
        filter(component_name == component$component_name[c]) %>%
        filter(project_name == project$project_name[p]) %>%
        mutate(ylab = factor(ylab_project, levels = rev(sort(ylab_project)))) %>%
        arrange(desc(adjusted_end), ylab) %>%
        mutate(task_id = paste0("task_", stringr::str_pad(row_number(), 2, pad = "0")))

      fig <- ggplot(figure_project, aes(x = adjusted_start, xend = adjusted_end,
                                       y = task_id, yend = task_id,
                                       color = performance_plan_element, alpha = work_intensity)) +
        geom_vline(aes(xintercept = Sys.Date()), col = "darkgray", linewidth = 1.5) +
        geom_segment(size = 5) +
        scale_color_manual(name = "Performance Plan",
                       values = c("Project lead" = "deeppink4",
                                  "Elem#1" = "chocolate1",
                                  "Elem#2" = "darkgoldenrod2",
                                  "Elem#3" = "darkolivegreen4",
                                  "Elem#4" = "cadetblue")) +
        scale_alpha_manual(name = "Work Intensity",
                       values = c("High" = 1,
                                  "Moderate" = 0.45,
                                  "Low" = 0.1,
                                  "Unknown" = 0)) +
        scale_x_date(limits = c(Sys.Date() - 31, Sys.Date() + 150),
             date_labels = "%Y-%m-%d",
             date_breaks ="1 month") +
        scale_y_discrete(labels = figure_project$ylab) +
        labs(title = project$project_name[p],
             x = "Date",
             y = "Role") +
        theme(plot.title.position = "plot") +
        theme(panel.grid.minor = element_blank()) +
        theme(axis.text.x = element_text(vjust = 0.5, hjust = 1)) +
        theme(legend.position = "top", legend.box = "vertical", legend.margin = margin())

      print(fig)
      cat("\n")
    }
  }
}
```

</details>

# Manuscripts

The following tables list the actions associated with each planned manuscript. Asterisks at the end of each role indicate project priority (one asterisk = mission critical work; two asterisks = high priority work).

<details>

<summary>Click to Expand</summary>

```{r manuscripts, results = "asis"}
for (m in 1:length(manuscripts)) {
  cat("\n\n## ", substr(manuscripts[m], 13, nchar(manuscripts[m])), "\n")

  manuscript_subset <- work_manuscripts %>%
    filter(project_name == manuscripts[m]) %>%
    select(-project_name) %>%
    rename("Role" = role_name,
           "Staff" = staff,
           "Planned End Date" = planned_end_date,
           "Status" = role_status)

  print(knitr::kable(manuscript_subset, format = "pipe", padding = 2))
}
```

</details>

# Reports

The following tables list the actions associated with each planned report.

<details>

<summary>Click to Expand</summary>

```{r reports, results = "asis"}
for (r in 1:length(reports)) {
   cat("\n\n## ", substr(reports[r], 10, nchar(reports[r])), "\n")

    report_subset <- work_reports %>%
      filter(project_name == reports[r]) %>%
      select(-project_name) %>%
      rename("Role" = role_name,
             "Staff" = staff,
             "Planned End Date" = planned_end_date,
             "Status" = role_status)

    print(knitr::kable(report_subset, format = "pipe", padding = 2))
  }
```

</details>

# Plan Elements (FY26)

The following tables list the roles for all PEP staff by performance plan element.

<details>

<summary>Click to Expand</summary>

```{r elements, results='asis'}
cat("::: {.panel-tabset}")
for (p in 1:length(pep)) {
  peep <- pep[p]
  cat("\n\n## ", peep, "\n")
  
  
  elements <- work_fy %>%
    filter(staff == peep) %>%
    select(pp) %>%
    unique() %>%
    arrange(pp) 
  
  for (e in 1:nrow(elements)) {

    element <- elements$pp[e]
    element_text <- paste0("---Element #", element, "---")
   
     cat("\n\n### ", 
        element_text
        , "\n")
    
    project_parents <- work_fy %>%
      filter(staff == peep,
             pp == element) %>%
      select(project_parents) %>%
      unique() %>%
      arrange(project_parents)
    
    for (c in 1:nrow(project_parents)) {
      project_parent <- project_parents$project_parents[c]
      cat("\n#### [", project_parent, "]{.ul}\n")
      
      projects <- work_fy %>%
      filter(staff == peep,
             pp == element,
             project_parents == project_parent) %>%
      select(project_name) %>%
      unique() %>%
      arrange(project_name)
      
      for (j in 1:nrow(projects)) {
        project <- projects$project_name[j]
        cat("\n##### .. ", project, "\n")
        
        if(element > 0) {
          roles <- work_fy %>%
            filter(staff == peep,
                   pp == element,
                   project_parents == project_parent, 
                   project_name == project,
                   role_name != 'Component lead') 
          
          if (nrow(roles) > 0) {
            roles <- roles %>%
              select(performance_plan_element, role_name) %>%
              unique() %>%
              arrange(performance_plan_element, role_name)
            
            for (r in 1:nrow(roles)) {
              cat("\n###### .. .. ", roles$role_name[r], "\n")
            }
          }
        }
      }
    }
  }
}
cat(":::")
```

</details>

# Prioritized Elements (FY26)

The following tables list the roles associated with Mission Critical or Top Priority projects for all PEP staff by performance plan element.

<details>

<summary>Click to Expand</summary>

```{r element_priorities, results='asis'}
cat("::: {.panel-tabset}")
for (p in 1:length(pep)) {
  peep <- pep[p]
  cat("\n\n## ", peep, "\n")
  
  
  elements <- work_fy_priorities %>%
    filter(staff == peep) %>%
    select(pp) %>%
    unique() %>%
    arrange(pp) 
  
  for (e in 1:nrow(elements)) {

    element <- elements$pp[e]
    element_text <- paste0("---Element #", element, "---")
   
     cat("\n\n### ", 
        element_text
        , "\n")
    
    project_parents <- work_fy_priorities %>%
      filter(staff == peep,
             pp == element) %>%
      select(project_parents) %>%
      unique() %>%
      arrange(project_parents)
    
    for (c in 1:nrow(project_parents)) {
      project_parent <- project_parents$project_parents[c]
      cat("\n#### [", project_parent, "]{.ul}\n")
      
      projects <- work_fy_priorities %>%
        filter(staff == peep,
             pp == element,
             project_parents == project_parent) %>%
      select(project_name) %>%
      unique() %>%
      arrange(project_name)
      
      for (j in 1:nrow(projects)) {
        project <- projects$project_name[j]
        cat("\n##### .. ", project, "\n")
        
        if(element > 0) {
          roles <- work_fy_priorities %>%
            filter(staff == peep,
                   pp == element,
                   project_parents == project_parent, 
                   project_name == project,
                   role_name != 'Component lead') 
          
          if (nrow(roles) > 0) {
            roles <- roles %>%
              select(performance_plan_element, role_name) %>%
              unique() %>%
              arrange(performance_plan_element, role_name)
            
            for (r in 1:nrow(roles)) {
              cat("\n###### .. .. ", roles$role_name[r], "\n")
            }
          }
        }
      }
    }
  }
}
cat(":::")
```

</details>

